---
title: "Using mcmcRanking package"
author: "Javlon Isomurodov"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4
)
```

## Preparing Data

```{r message=FALSE}
library(BioNet)
library(DLBCL)
library(genefilter)
library(impute)
```

First, we need to prepare network and p-values of genes. As an example, we took DLBCL (Diffuse large B-cell lymphoma expression data) data from [Bioconductor](https://bioconductor.org/packages/release/data/experiment/html/DLBCL.html).

```{r message=FALSE, results='hide'}
data(exprLym)
data(interactome)
network <- subNetwork(featureNames(exprLym), interactome)
network <- largestComp(network)
expressions <- impute.knn(exprs(exprLym))$data
t.test <- rowttests(expressions, fac = exprLym$Subgroup)
ttest.pval <- t.test[, "p.value"]
names(ttest.pval) <- rownames(expressions)
```

## Estimation of mixing time

We extract network and p-value for each gene as in [BioNet Tutorial](https://bioconductor.org/packages/release/bioc/html/BioNet.html). Next, transform network to `igraph` object, calculate likelihood of every gene with `FDR=1e-7` and run *MCMC* with parameter of iteration equal to `100000`. Function `sample_llh`, returns log-likelihoods of subgraph on every iteration.

```{r message=FALSE, fig.show="hold", warning=FALSE}
library(igraph)
library(mcmcRanking)
g <- graph_from_graphnel(network)
V(g)$pval <- ttest.pval[V(g)$name]
fb <- fitBumModel(V(g)$pval, F)
x <- exp(scoreFunction(fb, 1e-7))
deep <- repetition_depth(max(x))
V(g)$likelihood <- x^(1/2^deep)
y <- sample_llh(graph = g, module_size = 1, iter = 1e5, fixed_size = F)
plot(x=names(y), y = y, type = 'l',
     xlab = "iteration number", ylab = "log-likelihood")
```

Plot shows that log-liekelihood value stabilizes after about `20000` iterations. Thus we can estimate *MCMC* mxixing time as `20000` iterations.

## Calculating the probability

Next, run *MCMC* `1000` times and get probability of vertex as frequency of its presence  in a sampled subgraphs.

```{r message=FALSE, eval=FALSE}
iter <- 2e4
times <- 1e3
z <- mcmc_sample(graph = g, module_size = 1, times = times, iter = iter, fixed_size = F)
for(i in (deep-1):0){
  V(g)$likelihood <- x^(1/2^i)
  z <- mcmc_sample(graph = g, previous_mcmc = z, iter = iter, fixed_size = F)
}
p <- get_frequency(z)/times
r <- probabilistic_rank(graph = g, q = 1 - p)
```

Last two functions calculate probability of gene and connectivity preserving ranking respectivly. To increase the accuracy of the probability, increase the number of runs or parameter `times`.  
